A Survey of Quantitative Descriptions of Molecular Structure

### 2012-03-19 paper.tex
Some minor reformatting and text

A final class of descriptors are fingerprints. Traditionally, these
descriptors are represented in the form of bit strings. Binary
fingerprints can be divided into hashed fingerprints, where
substructures (such as paths of length $n$) are converted to a string
representation and then hashed into a randomly selected bit positions,
or keyed fingerprints, where each bit position corresponds to a unique
substructural feature. The nature of the features can range from
simple functional groups (hydroxyl, carbonyl) and topological
substructures (paths, chains, cycles) to atom environments
\cite{Bremser1978,Bender:2004aa} and pharmacophoric elements
\cite{Renner:2006aa}.

While the descriptors mentioned so far are derived from some form of
chemical structure, an alternative approach is to employ observed
biological activities of a molecule as descriptors of that molecule.
This approach was taken by Sedykh et al \cite{Sedykh:2011fk}   who
hypothesized that dose response data points from high throughput
dose-reponse assays could be employed as biological descriptors. They
observed that when such descriptors were combined with traditional
chemical structure descriptors, the predictive performance of models
developed using the combined set was better than those developed using
conventional approaches. A similar approach can be taken using the
PASS methodology \cite{Poroikov:2007aa}, though this is an indirect
method - one must predict the PASS profile from the chemical
structure, which can then be used in subsequent predictive models [REF?].

>>>>

\textbf{Fingerprints} A final class of descriptors are
fingerprints. Traditionally, these descriptors are represented in the
form of bit strings. Binary fingerprints can be divided into hashed
fingerprints, where substructures (such as paths of length $n$) are
converted to a string representation and then hashed into a randomly
selected bit positions, or keyed fingerprints, where each bit position
corresponds to a unique substructural feature. The nature of the
features can range from simple functional groups (hydroxyl, carbonyl)
and topological substructures (paths, chains, cycles) to atom
environments \cite{Bremser1978,Bender:2004aa} and pharmacophoric
elements \cite{Renner:2006aa}. Depending on the definition,
fingerprints can work with topology only or may require 3D
conformations. The latter class of fingerprints are usually related to
3D pharmacophores and most fingerprint definitions only require
connectivity information.

\textbf{Bioactivity as a descriptor} While the descriptors mentioned
so far are derived from some form of chemical structure, an
alternative approach is to employ observed biological activities of a
molecule as descriptors of that molecule.  This approach was taken by
Sedykh et al \cite{Sedykh:2011fk} who hypothesized that dose response
data points from high throughput dose-reponse assays could be employed
as biological descriptors. They observed that when such descriptors
were combined with traditional chemical structure descriptors, the
predictive performance of models developed using the combined set was
better than those developed using conventional approaches. A similar
approach can be taken using the PASS methodology
\cite{Poroikov:2007aa}, though this is an indirect method - one must
predict the PASS profile from the chemical structure, which can then
be used in subsequent predictive models [REF?].
